\newpage

\section{Histograms and Normal Distribution of Photocell Readings}
\label{s:photocell}

\subsection{Parts List}

\begin{enumerate}[itemsep=-5pt]
\item Laptop
\item CPX/CPB
\item USB Cable
\item Photocell
\item Resistor (10 kOhm)
\item Alligator Clips (x3)
\item Breadboard
\end{enumerate}

\subsection{Learning Objectives}
\begin{enumerate}[itemsep=-5pt]
\item Understand how a photocell responds to light level by measuring the voltage across the photocell in different light conditions
\item Learn how to create histograms of a noisy data signal
\item Understand mean and standard deviation and how that applies to Normal distributions.
\end{enumerate}

\subsection{Getting Started}

This lab is going to be similar to the potentiometer lab (See chapter \ref{s:voltage}). We are going to use a photocell though to vary the resistance instead of a potentiometer. Photocells are cool because they change their resistance solely based on the light intensity hitting the sensor rather than twisting a knob like the potentiometer.

\subsection{Assignment}

Turning in this assignment
Once you've done that upload a PDF with all of the photos and text below included. My recommendation is for you to create a Word document and insert all the photos and text into the document. Then export the Word document to a PDF. For videos I suggest uploading the videos to Google Drive, turn on link sharing and include a link in your PDF.

\subsubsection{Part 1}
\begin{enumerate}[itemsep=-5pt]
\item Include a video of you varying light conditions and watching the digital signal in the Plotter in Mu go up and down (make sure your face is in the video at some point and you state your name) - 50\%
\item Include your Python code in the appendix along with your data plotted in a Figure. Make sure to plot the voltage and not the digital output - 50\%
\end{enumerate}

\subsubsection{Part 2}
\begin{enumerate}[itemsep=-5pt]
\item Include a video of your circuit and explain how you captured the data to store on your computer (make sure your face is in the video at some point and you state your name) - 25\%
\item Include the mean, median, and standard deviation of all light levels in volts - 25\%
\item Include 3 histogram plots of low light, ambient light and high light levels. On top of the histogram I want you to plot the normal Gaussian distribution to see how close your histogram is to a Gaussian distribution  - 50\%
\end{enumerate}