\documentclass[11pt,a4paper,twoside,openright]{book}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{fullpage}
\usepackage[usenames,dvipsnames]{color}
%\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{datetime}
\usepackage{natbib} %REMOVE THIS IF YOU DON'T WANT BRACKETS ON YOUR PAPER
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{wrapfig}		%					allows wrapping text around figures
\usepackage{multicol}
%\usepackage{cite}
%\usepackage[labelfont=bf]{caption}
\usepackage{titlesec}
\titleformat{\section}[block]{\large\bfseries}{\thesection}{1em}{}
\usepackage{framed}
\usepackage{setspace}
\def\beq{\begin{equation}}
\def\eeq{\end{equation}}
\def\beqn{\begin{matrix}}
\def\eeqn{\end{matrix}}

\setlength{\marginparwidth}{0pt}

% Define commands to assure consistent treatment throughout document
\newcommand{\eqnref}[1]{(\ref{#1})}
 \newcommand{\class}[1]{\texttt{#1}}
 \newcommand{\package}[1]{\texttt{#1}}
 \newcommand{\file}[1]{\texttt{#1}}
 \newcommand{\BibTeX}{\textsc{Bib}\TeX}
\renewcommand{\thesection}{\Roman{section}.}
\renewcommand{\thesubsection}{\arabic{subsection}.}
\renewcommand{\thesubsubsection}{\alph{subsubsection}.}

\begin{document}

\begin{center}
\begin{LARGE}{\bf Aircraft Flight Mechanics}\end{LARGE}\\
\large
\vspace{22 mm}
\begin{singlespace}
   A Brief Textbook Presented to the \\ 
   Student Body of the University of South Alabama \\
\vspace{22 mm}
   by\\
\vspace{22 mm}
  Carlos Montalvo \\
\vspace{22 mm}
\end{singlespace}
{\itshape University of South Alabama}\\
{\bf Copyright $\copyright$ Carlos Jos\'{e} Montalvo}
\\ Last Update: \today
\end{center}

\newpage

\section*{Manuscript Changes}

\begin{enumerate}[itemsep=-5pt]
\item December 10th, 2020 - Moved manuscript to public Github repo
\item December 30th, 2020 - Added datetime to title page. Added
  section headers to RC aircraft design. Wrote text all the way
  through the airfoil selection. Added a references section.
\end{enumerate}

\newpage

\section*{Changes Needed}

\begin{enumerate}[itemsep=-5pt]
\item Need to finish section on how to design an RC aircraft
\item It would be nice to include some plots on RC aircraft design
  sections
\item I think an example RC aircraft design example would be good
\end{enumerate}

\tableofcontents

\newpage

\section{Particle Dynamics}
For this formulation we start with Newton's Second Law with no
approximations\cite{multibody}.
\begin{equation}
\sum\limits_{i=0}^N \vec{F}_{ji} = \frac{d\vec{p}_j}{dt}
\end{equation}
where $\vec{p}_j$ is the momentum of a particle. $\vec{F}_{ji}$ is a
force on the particle. The statement above states that sum of all
forces on a particle is equal to the time rate of change of
momentum. 

\subsection{Linear Dynamics for Systems of Particles}
If two particles are then considered the equation can
be written for both particles.
\begin{equation}
\begin{matrix}
\sum\limits_{i=0}^N \vec{F}_{1i} + \vec{f}_{12} = \frac{d\vec{p}_1}{dt} &
\sum\limits_{i=0}^N \vec{F}_{2i} + \vec{f}_{21} = \frac{d\vec{p}_2}{dt} 
\end{matrix}
\end{equation}
Note that the forces $\vec{f}_{12}$ and $\vec{f}_{21}$ are internal forces
experienced by each particle exerted on each other since they are
rigidly connected. Newton's Third Law states that for every action
there is an equal and opposite reaction. That is, $\vec{f}_{12} = -\vec{f}_{21}$. Thus, if both equations are added the following equation is
created
\begin{equation}
\sum\limits_{j=0}^P \sum\limits_{i=0}^N \vec{F}_{ji} =
\sum\limits_{j=0}^P \frac{d\vec{p}_j}{dt}
\end{equation}
where P is the number of particles. Typically the double summation
in F is written just as $\vec{F}$. 

\subsection{Rotational Dynamics for Systems of Particles}

Note that by construction, a system of particles rigidly connected can
now rotate about a center point. The center of mass of a system of
particles can be defined using the relationship below
\begin{equation}
\vec{r}_C = \frac{1}{m}\sum\limits_{j=0}^P m_j\vec{r}_{j}
\end{equation}
where
\begin{equation}
m = \sum\limits_{j=0}^P m_j
\end{equation}
This vector can then be used to create rotational dynamics starting
with the linear dynamics.
\begin{equation}
\sum\limits_{j=0}^P \sum\limits_{i=0}^N {\bf S}(\vec{r}_{Cj}) \vec{F}_{ji}  = \vec{M}_C = \sum\limits_{j=0}^P {\bf
  S}(\vec{r}_{Cj}) \frac{d\vec{p}_j}{dt}
\end{equation}
where ${\bf S}(\vec{r}_{Cj})$ is the skew symmetric matrix of the
vector from the center of mass to the jth particle which results in a cross product.
\section{Rigid Bodies}
At this point, many assumptions are made about the
system of particles.
\begin{enumerate}
\item The mass of each particle or rigid body is constant.
\item The Earth is assumed to be flat and not rotating. This allows
  the Earth to be used as an inertial frame.
\item The rigid body is not flexible and does not change shape. That
  is, the time rate of change of the magnitude of a vector
  $\vec{r}_{PQ}$ is zero for any arbitrary points P and Q attached to
  the rigid body.
\end{enumerate}
\subsection{Linear Dynamics}
Using all of these simplifications, the momentum term on the right can
be simplified to 
\begin{equation}
\sum\limits_{j=0}^P \vec{p}_j = m \vec{v}_{C/I}
\end{equation}
The derivation of the term above starts by deriving the position of
the center of mass as the following equation.
\begin{equation}
\vec{r}_{j} = \vec{r}_C + \vec{r}_{Cj}
\end{equation}
Taking one derivative results in the following equation
\begin{equation}
\vec{v}_{j/I} = \vec{v}_{C/I} + \frac{{}^Bd \vec{r}_{Cj}}{dt} +
{\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}
\end{equation}
where ${\bf S}(\vec{\omega}_{B/I})$ is the skew symmetric matrix of the
angular velocity vector which results in a cross product. This
equation comes from the derivative transport theorem. Since the body
is a rigid body the term $\frac{{}^Bd \vec{r}_{Cj}}{dt}=0$ resulting
in the equation below
\begin{equation}
\vec{v}_{j/I} = \vec{v}_{C/I} + {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}
\end{equation}
which any dynamicist knows as the equation for two points fixed on a
rigid body. This equation can then be substituted into the equation
for momentum such that.
\begin{equation}
\sum\limits_{j=0}^P \vec{p}_j =  \sum\limits_{j=0}^P m_j \left(\vec{v}_{C/I}
+ {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}\right)
\end{equation}
The first term reduces to 
\begin{equation}
\sum\limits_{j=0}^P m_j \vec{v}_{C/I} =  \vec{v}_{C/I}
\sum\limits_{j=0}^P m_j =  m \vec{v}_{C/I} 
\end{equation}
the second term reduces to zero since the sum of all particles from
the center of mass is by definition the center of mass and thus zero.
\begin{equation}
\sum\limits_{j=0}^P {\bf S}(\vec{\omega}_{B/I}) m_j\vec{r}_{Cj} =
{\bf S}(\vec{\omega}_{B/I})\sum\limits_{j=0}^P m_j\vec{r}_{Cj} = 0
\end{equation}
Plugging this result for momentum into Newton's equation of motion
yields. This is typically called Newton-Euler equations of motion.
\begin{equation}
\vec{F}_C = m \left(\frac{{}^Bd \vec{v}_{C/I}}{dt} +
{\bf S}(\vec{\omega})_{B/I} \vec{v}_{C/I} \right)
\end{equation}
\subsection{Rotational Dynamics}
Plugging in the expression for two points fixed on a rigid body
results in a much different expression. First let's expand the
rotational dynamic equations of particles using the assumptions made
for a rigid body.
\begin{equation}
\vec{M}_C = \frac{d}{dt}\sum\limits_{j=0}^P {\bf S}(\vec{r}_{Cj}) m_j\vec{v}_{j/I}
\end{equation}
Then the equation of two points fixed on a rigid body can be
introduced to obtain the following equation
\begin{equation}
\vec{M}_C = \frac{d}{dt}\sum\limits_{j=0}^P {\bf S}(\vec{r}_{Cj}) m_j\left(\vec{v}_{C/I} + {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}\right)
\end{equation}
expanding this into two terms yields
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left(\sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}) {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj} + \sum\limits_{j=0}^P {\bf S}(\vec{r}_{Cj})m_j\vec{v}_{C/I}\right)
\end{equation}
To simplify this further a useful equality is used for cross
products. That is ${\bf S}(\vec{a})\vec{b}=-{\bf
  S}(\vec{b})\vec{a}$. The equation above then changes to
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left(\left(-\sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}){\bf
  S}(\vec{r}_{Cj})\right)\vec{\omega}_{B/I} - {\bf S}(\vec{v}_{C/I})
\sum\limits_{j=0}^P \vec{r}_{Cj}m_j\right)
\end{equation}
Notice, that parentheses were placed around the first term to isolate
the angular velocity. This is because the angular velocity is constant
across the system of particles. The term on the right has also been
altered slightly to isolate the fact that the velocity of the center
of mass is independent of the system of particles. With the equation
in this form it is easy to see that the term on the right is zero
because it is the definition of the center of mass. The equation then
reduces to 
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left(\sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}){\bf
  S}(\vec{r}_{Cj})^T\right)\vec{\omega}_{B/I} 
\end{equation}
Notice again that minus sign has been removed. The skew symmetric
matrix has an interesting property where the transpose is equal to the
negative of the original matrix. The term in brackets is a well known
value for rigid bodies and is known as the moment of inertia for rigid
bodies. 
\begin{equation}
{\bf I}_C =  \sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}){\bf
  S}(\vec{r}_{Cj})^T
\end{equation}
This results in the kinematic equations of motion for rigid bodies to
the simple equation below.
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left({\bf I}_C\vec{\omega}_{B/I} \right)
\end{equation}
With the equation in this form it is finally possible to carry out the
derivative
\begin{equation}
\vec{M}_C = \frac{{}^Bd ({\bf I}_C\vec{\omega}_{B/I})}{dt} + {\bf
  S}(\vec{\omega}_{B/I}){\bf I}_C\vec{\omega}_{B/I}
\end{equation}
The first term requires the chain rule to perform the derivative
however the body frame derivative of the moment of inertia matrix is
simply zero. Therefore the equation can simply be written as
\begin{equation}
\vec{M}_C = {\bf I}_C\frac{{}^Bd (\vec{\omega}_{B/I})}{dt} + {\bf
  S}(\vec{\omega}_{B/I}){\bf I}_C\vec{\omega}_{B/I}
\end{equation}
\section{Aircraft Convention}
Aircraft convention involves using the Newton-Euler equations of
motion to describe the aircraft\cite{etkins}. Typically the position of the
aircraft is written as 
\begin{equation}
{\bf C}_I(\vec{r}_C) = \begin{Bmatrix} x \\ y \\ z \end{Bmatrix}
\end{equation}
The derivative of the position vector is the velocity vector is then
written as
\begin{equation}
{\bf C}_I(\vec{v}_{C/I}) = \begin{Bmatrix} \dot{x} \\ \dot{y} \\ \dot{z} \end{Bmatrix}
\end{equation}
However, body frame coordinates are typically used to
describe the velocity vector such that
\begin{equation}
{\bf C}_B(\vec{v}_{C/I}) = \begin{Bmatrix} u \\ v \\ w \end{Bmatrix}
\end{equation}
In order to relate the body frame components of the velocity vector
the inertial frame coordinates a transformation matrix is used. The
transformation from the inertial frame to the body frame involves
three unique rotations. The first is a rotation about the z-axis such
that
\begin{equation}
{\bf C}_A(\vec{v}_{C/I}) = \begin{bmatrix} cos(\psi) & sin(\psi) & 0
  \\ -sin(\psi) & cos(\psi) & 0 \\ 0 & 0 & 1 \end{bmatrix} {\bf C}_I(\vec{v}_{C/I})
\end{equation}
this rotation is called the yaw or heading rotation. From here the
intermediate frame is rotated about the y-axis such that
\begin{equation}
{\bf C}_{NR}(\vec{v}_{C/I}) = \begin{bmatrix} cos(\theta) & 0 &
  -sin(\theta) \\ 0 & 1 & 0 \\ sin(\theta) & 0 & cos(\theta) \end{bmatrix} {\bf C}_A(\vec{v}_{C/I})
\end{equation}
this rotation is called the pitch angle rotation. Finally the no roll
frame is rotated through the x-axis such that
\begin{equation}
{\bf C}_{B}(\vec{v}_{C/I}) = \begin{bmatrix} 1 & 0 & 0 \\ 0 & cos(\phi) & sin(\phi)
  \\ 0 & -sin(\phi) & cos(\phi) \end{bmatrix} {\bf C}_{NR}(\vec{v}_{C/I})
\end{equation}
Putting all the equations together yields
\begin{equation}\label{e:xyzdot}
\begin{Bmatrix} \dot{x} \\ \dot{y} \\ \dot{z}   \end{Bmatrix} = [\textbf{T}_{IB}]
\begin{Bmatrix} u \\ v \\ w \end{Bmatrix}
\end{equation}
where 
\begin{equation}\label{e:TIB}
\textbf{T}_{IB} = \begin{bmatrix} c_{\theta}c_{\psi} &
s_{\phi}s_{\theta}c_{\psi}-c_{\phi}s_{\psi} &
c_{\phi}s_{\theta}c_{\psi} + s_{\phi}s_{\psi} \\ c_{\theta}s_{\psi} &
s_{\phi}s_{\theta}s_{\psi} + c_{\phi}c_{\psi} &
c_{\phi}s_{\theta}s_{\psi} - s_{\phi}c_{\psi} \\
-s_{\theta} & s_{\phi}c_{\theta} & c_{\phi}c_{\theta}
\end{bmatrix}
\end{equation}
Standard shorthand notation is used for trigonometric
functions: $ cos(\alpha) \equiv c_{\alpha} $ , $ sin(\alpha) \equiv
s_{\alpha} $ , and $ tan(\alpha) \equiv t_{\alpha} $. These three
angles are known as the standard Euler angle rotation sequence of an
aircraft in free flight. The angular
velocity of a body is typically written as 
\begin{equation}
{\bf C}_B(\vec{\omega}_{B/I}) = \begin{Bmatrix} p \\ q
  \\ r \end{Bmatrix} = p \hat{I}_B + q \hat{J}_B + r \hat{K}_B
\end{equation}
There are no inertial components for the angular velocity
vector. However, a relationship can be derived relating the
derivatives of the Euler angles. The angular velocity can be written
in vector form such that
\begin{equation}
\vec{\omega}_{B/I} = \dot{\psi} \hat{K}_A + \dot{\theta} \hat{J}_{NR} +
\dot{\phi} \hat{I}_B
\end{equation}
relating the unit vectors $\hat{K}_A$ and $\hat{J}_{NR}$ to the body
frame using the planar rotation matrices results in the equation
below. Note that NR is denoted as the ``No-Roll" frame.
\begin{equation}\label{e:ptpdot}
\begin{Bmatrix} \dot{\phi} \\ \dot{\theta} \\ \dot{\psi} \end{Bmatrix}
= [\textbf{H}]
\begin{Bmatrix} p\\q\\r\end{Bmatrix}
\end{equation}
where
\begin{equation}
\textbf{H}=\begin{bmatrix} 1 & s_{\phi}t_{\theta} & c_{\phi}t_{\theta} \\ 0 &
c_{\phi} & -s_{\phi} \\ 0 & s_{\phi}/c_{\theta} &
c_{\phi}/c_{\theta} \end{bmatrix}
\end{equation}
with all this information the Newton-Euler equations of motion can be
used to form the equation below.
\begin{equation}\label{e:uvwdot} 
\begin{Bmatrix} \dot{u} \\ \dot{v} \\ \dot{w} \end{Bmatrix} = 
\frac{1}m \begin{Bmatrix} X\\Y\\Z\end{Bmatrix}-\begin{bmatrix} 0 & -r
& q \\ r & 0 & -p \\ -q & p & 0 \end{bmatrix} \begin{Bmatrix} u\\v\\w\end{Bmatrix}
\end{equation}
Using the rotational dynamic equations for rigid bodies, the equation
for the derivative of angular velocity can be found as
\begin{equation}\label{e:pqrdot} 
\begin{Bmatrix} \dot{p} \\ \dot{q} \\ \dot{r} \end{Bmatrix} = {\bf I}_C^{-1}\left(
\begin{Bmatrix} L\\M\\N\end{Bmatrix}-\begin{bmatrix} 0 & -r
& q \\ r & 0 & -p \\ -q & p & 0 \end{bmatrix} {\bf I}_C\begin{Bmatrix} p\\q\\r\end{Bmatrix}\right)
\end{equation}
Note that standard aircraft forces and moments are applied to the
body. The forces are typically written as X,Y and Z while the moments
are given as L,M and N. They can be written in component form using
the equations below.
\begin{equation}
{\bf C}_B(\vec{F}_{C}) = \begin{Bmatrix} X \\ Y
  \\ Z \end{Bmatrix} = X \hat{I}_B + Y \hat{J}_B + Z \hat{K}_B
\end{equation}
\begin{equation}
{\bf C}_B(\vec{M}_{C}) = \begin{Bmatrix} L \\ M
  \\ N \end{Bmatrix} = L \hat{I}_B + M \hat{J}_B + N \hat{K}_B
\end{equation}

\section{Forces on Aircraft}
To form the sum of the forces on an aircraft the assumptions are made
that only gravity and aerodynamic forces act. For rockets, a
propulsion model can be added. Still the gravitational force is shown
below. 
\subsection{Gravity}
The weight contribution is assumed to be a constant force applied to
the aircraft. The equation below is the gravitational force applied in
the body frame.
\begin{equation}\label{e:wforce}
\begin{Bmatrix} X_W \\ Y_W \\ Z_W \end{Bmatrix} = mg \begin{Bmatrix}
-s_{\theta} \\ s_{\phi}c_{\theta} \\ c_{\phi}c_{\theta} \end{Bmatrix}
\end{equation}
\subsection{Aerodynamics}\label{s:aerodynamics}
Aircraft aerodynamics are written using a taylor series expansion
about a trim point\cite{Phillips,AndersonD}. That is, the aerodynamic forces are given by
\begin{equation}
\vec{F} = \vec{F}_0 + \frac{\partial \vec{F}}{\partial \vec{x}}(\vec{x}-\vec{x}_0)
\end{equation}
where $\vec{x} = [x,y,z,\phi,\theta,\psi,u,v,w,p,q,r]^T$. The partial
derivative is thus expanded such that
\begin{equation}
\frac{\partial \vec{F}}{\partial \vec{x}} = \begin{bmatrix} \frac{\partial
    \vec{F}}{\partial x} & \frac{\partial \vec{F}}{\partial y} & ... &
  \frac{\partial \vec{F}}{\partial r} \end{bmatrix}
\end{equation}
To find all of the partial derivative the forces are first written
using a combination of dynamic pressure and coefficients that are
functions of geometry and Reynolds number rather than speed, pressure
and size. A general lift force can be written using the equation below
\begin{equation}
L = \frac{1}2\rho {V_{\infty}}^2 S C_L
\end{equation}
where $\rho$ is the atmospheric density, $V_{\infty}$ is the
free-stream velocity, S is the planform area of the wing and $C_L$ is
the lift coefficient. 
\begin{equation}\label{e:vtotal}
  V_{\infty} = \sqrt{{u_a}^2 + {v_a}^2 + {w_a}^2}
\end{equation}
The subscript 'a' above denotes the velocity of the aircraft plus the
atmospheric disturbance. 
\begin{equation}\label{e:atm}
\begin{Bmatrix} u_a \\ v_a \\ w_a \end{Bmatrix} =
\begin{Bmatrix} u \\ v \\ w \end{Bmatrix} +
{\textbf{T}^T}_{IB} \begin{Bmatrix}
 V_x \\ V_y \\ V_z \end{Bmatrix}
\end{equation}
Note that the dynamic pressure is different for
a rocket or projectile. A similar expression can be created for a
generic moment such that
\begin{equation}
M = \frac{1}2\rho {V_{\infty}}^2 S \bar{c} C_M
\end{equation}
where $\bar{c}$ is the mean chord of the aircraft. The dynamic pressure $q_{\infty} =
\frac{1}2\rho {V_{\infty}}^2 S$ can be used to non-dimensionalize the forces, thus $L/q_{\infty} = C_L$. This means
that the equation involving partial derivatives can be written as
\begin{equation}
\frac{\partial \vec{C}_F}{\partial \vec{x}} = \begin{bmatrix} \frac{\partial
    \vec{C}_F}{\partial x} & \frac{\partial \vec{C}_F}{\partial y} & ... &
  \frac{\partial \vec{C}_F}{\partial r} \end{bmatrix}
\end{equation}
If the vector is then expanded to include the components of the vector
$\vec{F}$ the partial derivatives expand to
\begin{equation}
\frac{\partial \vec{C}_F}{\partial \vec{x}} = \begin{bmatrix}
  \frac{\partial C_X}{\partial x} & \frac{\partial C_X}{\partial y} &
  ... & \frac{\partial C_x}{\partial r} \\ \frac{\partial C_Y}{\partial x} & \frac{\partial C_Y}{\partial y} &
  ... & \frac{\partial C_Y}{\partial r} \\ \frac{\partial C_Z}{\partial x} & \frac{\partial C_Z}{\partial y} &
  ... & \frac{\partial C_Z}{\partial r} \end{bmatrix}
\end{equation}
shorthand can be adopted for the forces above such that
$\frac{\partial C_Y}{\partial x} = C_{Yx}$. Using this shorthand the
equation above can be written as.
\begin{equation}
\frac{\partial \vec{C}_F}{\partial \vec{x}} = \begin{bmatrix}
  C_{Xx} & C_{Xy} &
  ... & C_{Xr} \\ C_{Yx} & C_{Yy} &
  ... & C_{Yr} \\ C_{Zx} & C_{Zy} &
  ... & C_{Zr} \end{bmatrix}
\end{equation}
The coefficients listed above are standard coefficients that all
aircraft have. A similar matrix can be formulated for the moments on
an aircraft. When system identifying an aircraft all of these
coefficients may be determined. However, many of these terms are
zero. For example, all coefficients with respect to x y and z are
zero. That is, $C_{Xx} = C_{Yx} = ... C_{Nx} = C_{Xy} = ... C_{Nz} =
0$. Other coefficients can be set to zero as well. 
\subsubsection{Aircraft Aerodynamics}
For aircraft, some further simplifications are made. Some of the
coefficients defined above are combined to be written as functions of the
angle of attack($\alpha$) and sideslip($\beta$).
\begin{equation}\label{e:aoa}
\alpha = tan^{-1}\left(\frac{w_a} {u_a} \right)
\end{equation}
\begin{equation}\label{e:beta}
\beta = sin^{-1}\left(\frac{v_a}{V_{\infty}}\right)
\end{equation}
Transforming the equations into these formulations gives rise to
coefficients such as $C_{L\alpha}$ which is the change in lift as a
function of angle of attack and $C_{Y\beta}$ which is the change in
Y-Force as a function of sideslip. Using all of the coefficients
defined above taking into account the change to lift and drag, the body
aerodynamic force is calculated using the equation below.
\begin{equation}\label{e:aforce}
\begin{Bmatrix} X_A \\ Y_A \\ Z_A \end{Bmatrix} = \frac{1}2\rho
{V_{\infty}}^2 S \begin{Bmatrix}
C_Ls_{\alpha}-C_Dc_{\alpha}+C_{x_{\delta_t}} \delta_t
\\ C_{y{\beta}}{\beta}+C_{y{\delta}_r}{\delta}_r+C_{yp}\frac{pb}{2V_{\infty}}
+ C_{yr}\frac{rb}{2V_{\infty}} \\ -C_Lc_{\alpha}-C_Ds_{\alpha} \end{Bmatrix}
\end{equation}
Where the lift and drag coefficients are:
\begin{equation}\label{e:liftdrag}
\begin{Bmatrix} C_L \\ C_D \end{Bmatrix} = \begin{Bmatrix} C_{L0} +
C_{L\alpha}\alpha + C_{Lq}\frac{q\bar{c}}{2V_{\infty}} + C_{L\delta_e}\delta_e \\   C_{D0} +
C_{D\alpha}\alpha^2 \end{Bmatrix}
\end{equation}
The body aerodynamic moment is also computed using an aerodynamic expansion.
\begin{equation}\label{e:LMN}
\begin{Bmatrix} L_A \\ M_A \\ N_A \end{Bmatrix} = \frac{1}2\rho
{V_{\infty}}^2 S \bar{c} \begin{Bmatrix} C_{l\beta}\beta + C_{lp}\frac{pb}{2V_{\infty}} + C_{lr}\frac{rb}{2V_{\infty}} + C_{l\delta_a}{\delta_a} + C_{l\delta_r}{\delta_r}
\\  C_{m0} + C_{m\alpha}\alpha + C_{mq}\frac{q\bar{c}}{2V_{\infty}}+ C_{m\delta_e}\delta_e  \\ C_{np}\frac{pb}{2V_{\infty}} + C_{n\beta}\beta + C_{nr}\frac{rb}{2V_{\infty}} + C_{n\delta_a}\delta_a + C_{n\delta_r}\delta_r \end{Bmatrix}
\end{equation}
The aerodynamic coefficients in equations
(\ref{e:aforce}), (\ref{e:liftdrag}) and (\ref{e:LMN}) can be obtained from
flight data, aerodynamic modeling and windtunnel tests. Notice that
the only coefficients remaining are coefficients from angle of attack,
sideslip and angular velocities. Furthermore, the coefficients for
angular velocities are also non-dimensionalized by terms such as
$b/(2V_{\infty})$ where $b$ is the wingspan of the aircraft and
$\bar{c}$ is the mean chord of the aircraft. These terms are
introduced to fully non-dimensionalize the coefficients. Notice, as
well that four extra terms were also introduced. These will 
be discussed in more detail in the control section however the four
terms are the aileron control surface $\delta_a$, the elevator
control surface $\delta_e$, the rudder control surface $\delta_r$ and
the thrust control value $\delta_t$. 

\subsubsection{Projectile Aerodynamics}

To fully define the projectile aerodynamics some more assumptions are
made about the projectile.
\begin{enumerate}
\item The projectile is axially symmetric
\item The aerodynamic forces are not necessarily formulated at the center
  of mass
\item The projectile has the potential to be spinning rapidly thus
  interacting with the surrounding atmosphere
\end{enumerate}
For a projectile the dynamic pressure is written as
\begin{equation}
Q = \frac{\pi}{8}\rho V_{\infty}^2d^2 
\end{equation}
The aerodynamic forces on the projectile are modeled using taylor series
ballistic expansions with known coefficients similar to the aircraft
model only slightly different assumptions are made given the dynamics
of the projectile. The subscripts in the equation below stand for
steady and unsteady aerodynamics.
\begin{equation}\label{e:aeroF}
\begin{Bmatrix} X_A \\ Y_A\\ Z_A \end{Bmatrix} = \begin{Bmatrix}
  X_{SA} \\ Y_{SA}\\ Z_{SA} \end{Bmatrix} + \begin{Bmatrix}
  X_{UA} \\ Y_{UA}\\ Z_{UA} \end{Bmatrix} = 
Q \begin{Bmatrix} -C_{X_0} - C_{X_2}\frac{v^2+w^2}{V^2}\\
-C_{Y_\beta}\frac{v}{V}\\
-C_{N_\alpha}\frac{w}{V}
\end{Bmatrix} + Q \begin{Bmatrix} 0
  \\ C_{Y_{p\alpha}}\frac{w}{V}\frac{pd}{2V} \\ C_{Z_{p\alpha}}\frac{v}{V}\frac{pd}{2V}\end{Bmatrix}
\end{equation}
In this equation, $Q$ is the dynamic pressure, $d$ is the aerodynamic
reference area, $C_{X_0}$ is the zero-yaw axial force
coefficient,$C_{X_2}$ is the yaw-squared axial force coefficient,
$C_{N_\alpha}$ is the normal force derivative coefficient,
$C_{Y_{p\alpha}}$ is the Magnus force coefficient, and
$V=\sqrt{u^2+v^2+w^2}$ is the total velocity of the projectile. 
The aerodynamic moments acting on the projectile are the pitching,
pitch damping, Magnus, and roll damping moments. Pitching and Magnus
moments are given by taking the cross product of the normal and Magnus
forces given in (\ref{e:aeroF})  with the position vector from the
center of mass to the center of pressure and location of Magnus force,
respectively. The total aerodynamic moments are given in
Eqn. (\ref{e:aeroM}). 
\begin{equation}\label{e:aeroM}
\begin{Bmatrix} L_A \\ M_A \\ N_A \end{Bmatrix}=
{\bf S}_B(\vec{r}_{CG,COP})\begin{Bmatrix}
  X_{SA} \\ Y_{SA}\\ Z_{SA} \end{Bmatrix} + {\bf S}_B(\vec{r}_{CG,MCOP})\begin{Bmatrix}
  X_{UA} \\ Y_{UA}\\ Z_{UA} \end{Bmatrix} + 
Qd \begin{Bmatrix}
C_{l_p} \frac{pd}{2V} \\
C_{m_q} \frac{qd}{2V}\\
C_{n_r} \frac{rd}{2V} \end{Bmatrix}
\end{equation}
Here, ${\bf S}_B(\vec{r}_{CG,COP})$ is the skew-symmetric operator
acting on the position vector from the center of mass to the center of
pressure expressed in the projectile body frame. Furthermore, ${\bf S}_B(\vec{r}_{CG,MCOP})$ is the skew-symmetric operator
acting on the position vector from the center of mass to the magnus
center of pressure expressed in the projectile body frame. Typically
the center of mass is defined from the rear of the projectile such
that 
\begin{equation}
{\bf C}_B(\vec{r}_{CG}) = \begin{Bmatrix} SL_{CG} \\ BL_{CG} \\ WL_{CG} \end{Bmatrix}
\end{equation}
Similarly, the center of pressure is defined from the rear of the
projectile such that
\begin{equation}
{\bf C}_B(\vec{r}_{COP}) = \begin{Bmatrix} SL_{COP} \\ BL_{COP} \\ WL_{COP} \end{Bmatrix}
\end{equation}
The vector $\vec{r}_{CG,COP}$ is then simply the different between
both vectors.
\begin{equation}
\vec{r}_{CG,COP} = \vec{r}_{COP}-\vec{r}_{CG}
\end{equation}
The damping coefficient defined in equation (\ref{e:aeroM})
include $C_{l_p}$ which is the roll damping coefficient while $C_{m_q}$ is the pitch damping
coefficient. These coefficients are added which essentially inhibit
angular motion of the projectile. In addition, to these coefficients,
sometimes magnus coefficients are given as pure moments rather 
than forces acting at a distance. This can be given in the equation
below. 
\begin{equation}
M_{UA} = Qd (-C_{M\alpha}\frac{v}{V} + C_{N_{p\alpha}}\frac{w}{V}\frac{pd}{2V})
\end{equation}
Where $C_{M\alpha}$ replaces the moment produced by $C_{N\alpha}$ and
$C_{N_{p\alpha}}$ replaces the moment produced by
$C_{Y_{p\alpha}}$. It is possible to derive an equation between the
two different representations as given by the equations below.

\begin{equation}
\begin{matrix}
C_{M\alpha} = \frac{(SL_{COP}-SL_{CG})C_{N\alpha}}{d} \\
C_{N_{p\alpha}} = \frac{(SL_{MAG}-SL_{CG})C_{Y_{p\alpha}}}{d}
\end{matrix}
\end{equation}

\section{Stability and Control}

Controllability is formally stated as a system where any initial
state $x(0)=x_0$ and final state $x_1,t_1>0$, there exists a piecewise
continuous input $u(t)$ such that $x(t_1)=x_1$. 
For a fixed wing aircraft the system has 12 states with 8 dynamic
modes and 4 zero or rigid body modes. For a fixed wing aircraft the system has 12 states with 8 dynamic
modes and 4 zero or rigid body modes. A conventional aircraft has 4
controls to control these 12 modes. The easiest way to test the 
controllability of a system  is to compute the
controllability matrix. However, the controllability matrix must be
computed using a linearized model such that
$\dot{\vec{x}}=A\vec{x}+B\vec{u}$. In order to do this the aircraft
must be in equilibrium. For this example the aircraft is
set with an initial velocity of $20~m/s$ at an altitude of
$200~m$. The altitude command is set to $200~m$ and the heading
command is set to zero. Given the zero heading angle command and the
symmetry of the configurations investigated the rudder and aileron
commands are set to zero. Thus, only the thrust and elevator controls
are activated for the trimming procedure. Each configuration is
simulated for 200 seconds or until the derivatives of all states
except $\dot{x}$ are within a required tolerance. Using this
equilibrium point a linear model can be computed by using forward
finite differencing assuming that the
aircraft model is put in the form $\dot{\vec{x}} = F(\vec{x},\vec{u})$.
\begin{equation}
\dot{\vec{\delta x}} = \frac{F(\vec{x_0}+\Delta \vec{x_0},\vec{u_0})-F(\vec{x_0},\vec{u_0})}{\Delta
  \vec{x}}\vec{\delta x} + \frac{F(\vec{x_0},\vec{u_0}+\Delta
  \vec{u})-F(\vec{x_0},\vec{u_0})}{\Delta \vec{u}}\vec{\delta u}
\end{equation}
This linear model is the classic linear model where
$\dot{\vec{\delta x}}=A\vec{\delta{x}}+B\vec{\delta{u}}$. Using this linear model, the
controllability matrix can be computed as
\begin{equation}
W_C = [B~AB~A^2B~A^3B~...~A^{N-1}B]
\end{equation}
where N is the number of states in the system. With the controllability
matrix formulated, the rank of the matrix is computed. If the
$rank(W_C)=N$ the system is said to be controllable.

\section{PID Control}

For a conventional PID controller of an aircraft, the rudder, elevator
and aileron commands are set to 
\begin{equation}
\begin{matrix}
\delta_{r} =-K_v v \\
\delta_{e} = K_p(\theta-\theta_{c})+K_d{\dot \theta} \\ 
\delta_a = K_p(\phi-\phi_{c})+K_d{\dot \phi}
\end{matrix}
\end{equation}
The Euler angle commands $\phi_{c}$ and $\theta_{c}$ are set using the
following relationships:
\begin{equation}
  \begin{matrix}
  \phi_{c} = K_p(\psi-\psi_c)+K_d\dot{\psi} \\
  \theta_{c} = K_p(z-z_c) + K_d\dot{z} + K_{I}\int{z-z_c}dt
  \end{matrix}
\end{equation}
The control scheme defined above is a conventional inner loop-outer
loop control of a fixed wing aircraft using a PID tracking
controller.

\section{Radio Controlled Aircraft Design}

Many principles for large aircraft design can be applied to smaller
radio controlled aircraft but understand that many of the aerodynamic
principles are not quite well defined for slow and small
aircraft. These aircraft have low Reynolds number which often exhibits
odd phenomena. For example, airfoil selection is really not so much a
design point other than ease of manufacturing rather than maximizing
lift coefficient. Remember that Reynolds number can be defined from
the equation below where $\rho$ is the density at sea-level (1.225
kg/$m^3$ in SI units), $V$ is
the desired flight speed, $\bar{c}$ is the mean aerodynamic chord and
$\mu_{\infty}$ is the viscosity of air which in SI units is 1.81e-5
$kg/(m-s)$. 

\beq
Re = \frac{\rho V \bar{c}}{\mu_{\infty}}
\eeq

You can tell that flying a small aircraft ($\bar{c}$) and flying slow
$(V)$ results in a low Reynolds number. Either way the procedure below
has produced some great aircraft and the tools you'll learn along the
way will help you in your future aerospace engineering career. If
you're not an engineer then this text will at least give you an
appreciation for what goes into aircraft design. If you'd much rather
watch youtube videos than read this document, feel free to watch my
Youtube playlist on radio controlled aircraft design\cite{RCYoutube}

\subsection{Vehicle Type Selection and Requirements}

In the very beginning of your design you need to decide on the type of
aircraft you want to build. Designing a glider versus an aerobatic
airplane will result in vastly different engineering design
decisions. For example, a glider is going to have very long slender
wings while an aerobatic airplane is going to have somewhat shorter
wings with large control surfaces. I suggest you select from the
following types of aircraft and then move on: Gliders, Trainers, Sport
Aerobatic, Racers. If you'd like to build a scale aircraft there isn't
much to design since the shape of the aircraft is pretty much
built. If you do go with a scale aircraft this textbook isn't really
for you since you aren't really building a scratch build
aircraft. You're more just copying someone else's design. In that case
you may as well just buy a kit or watch some videos on balsa wood
construction and an overview of all the electronics required for RC
aircraft flight.

If you selected one of the other styles you're ready to move onto to
the next stage which is requirements. There is so much literature on
Systems Engineering, top level requirements, functional
requirements and derived requirements. The bottom line is you need to
determine what you want your aircraft to do. Do you want it to fly
straight up, upside down? Do you want the aircraft to be hand
launched? Land on a runway? Determine what you want the aircraft to do
and create a bulleted list of those requirements. Throughout the
design you can refer to these requirements and make sure you are
satisfying these requirements. If this is your first build then you
may just have one requirement and that is to take off and land without
crashing. But think a bit deeper. Do you want to turn the vehicle? Do
you want full channel control for roll, pitch and yaw or just yaw
control? Do you want landing gear? What sort of flying characteristics
do you want? Be as specific as possible here. 

\subsection{Initial Design - Hand Sketch and Aspect Ratio}

Once you have an idea of what the aircraft type is and what the
requirements are it's time to hand sketch your aircraft. Try and use
engineering paper, french curves, a ruler and a compass. Make this
hand sketch look nice so you can use it in your future design. Draw
your sketch to scale. You might be wondering, ``how do I draw the
aircraft without knowing what my wing loading or thrust to weight
ratio is?". The answer comes from my late aircraft design professor
Dr. Mikolowski (RIP). He would always say ``If it looks good, it flies
good". After designing so many aircraft and seeing so many scratch
builds from my students I can honestly say that this is true
100\%. If you're reading this now it means that there is already over
100 years of aircraft technology on the internet for you to research
and see what other aircraft look like. Make your aircraft look like
that but make sure it fits into your aircraft type and make sure it
satisfies your requirements from above. If one of your requirements
was to hand launch, then make sure your drawing reflects a vehicle
without landing gear and a place to grab the aircraft. If you wanted
full channel support make sure to include all the control
surfaces. Think about where you want the propulsion system to go and
how you're going to access the electronics before you fly. Think about
what you want the wing to look like. Make it as big as you think it
needs to fly. Use your intuition. This is an art. So much of
engineering is an art.

Once your aircraft sketch is complete (make sure to do a front view,
side view and top view of your aircraft), it's time to take down some
wing characteristics. This is why you need to draw your sketch to
scale. Measure the length of the wing (wingspan $b$) and the chord at the
root ($c_r$) and the tip ($c_t$). Compute the area of the wing ($S$) using
the area of a trapezoid or rectangle depending on the shape of your
wing. Once you have the wingspan and area you can compute the aspect
ratio of your aircraft.

\beq
AR = \frac{b^2}{S}
\eeq

The general rule is that the larger the aspect ratio the more
aerodynamically efficient your aircraft will be. This is why gliders
have very long and slender wings. At the same time, high aspect ratio
wings suffer from larger bending moments and can flex considerably in
flight. Finally, it's important to compute the mean aerodynamic chord
of the wing. This is basically the average chord of your wing. If you
create a rectangular wing your mean aerodynamic chord is just the
chord of the wing since it's constant. If not you'll need to integrate
over the length of the wing using the formula below where y = 0 is the
centerline of the vehicle and y=b/2 is the wingtip on the right
side\cite{caughey}. 

\beq
\bar{c} = \frac{2}{S}\int^{b/2}_0 c^2 dy
\eeq

\subsection{Weight Estimate - Tabular Approach}

Once you have an idea of the overall shape it's time to estimate how
heavy the aircraft will be. First think about the fundamental
components of an aircraft. If you're not familiar with any of the
components below, just type the item into Google and you'll find
numerous articles and Youtube Videos about each component.

\begin{enumerate}[itemsep=-5pt]
\item ESC - Electronic Speed Controller
\item Battery - Assume for now that you'll be using a 1500mAh 3S 
  Battery unless you're building a micro aircraft in which case you
  might end up using a 600 mAh 2S or even a 300 mAh 1S. Think about
  the size of your aircraft. You will do more sophisticated battery
  design in the future.
\item Motor and Propeller - Again select something that is in the
  ballpark of the aircraft you're building. You'll do a redesign later.
\item Servos
\item Receiver
\item Control Linkages and Servo Horns
\item Fuselage
\item Main Wing
\item Tail both Horizontal and Vertical
\item Payload
\end{enumerate}

For each of the components above you need to estimate the weight of
these components. The only way to do that is to either look up the
weight of similar aircraft to the one you're designing or find
components that you think will work for your aircraft and add up all
of the weights. The most difficult part is going to be estimating the
empty weight of the aircraft which is just the structure of the
aircraft. For these estimates you need to decide what materials you
plan on using. Is your aircraft made out of foam, balsa, carbon fiber
or some type of combination. Perform an initial material selection and
then use that to estimate your weight. Create a table in a spreadsheet
type program or a numerical computer program so that you can go back
and change weights as your design progresses. Use the spreadsheet or
numerical program to compute the total weight of your aircraft. This
is your maximum takeoff weight.

\subsection{Airfoil Selection and 2D and 3D Lift}

The section \ref{s:aerodynamics} details the aerodynamic forces on the
aircraft. In this stage of design we will only be looking at the lift
equation.
\begin{equation}
L = \frac{1}2\rho {V}^2 S C_L
\end{equation}
The coefficient of lift $C_L$ is the lift coefficient of the
aircraft. This coefficient is a function of Reynolds number, angle of
attack, airfoil shape and wing shape. Recall that the angle of attack
is the angle between the zero lift line of the airfoil and the free
stream air. Breaking the velocity vector into components yields the
equation below where $w$ is the airspeed along the z-axis of the
vehicle and $u$ is the velocity along the x-axis. 
\begin{equation}\label{e:aoa}
\alpha = tan^{-1}\left(\frac{w} {u} \right)
\end{equation}
Using the angle of attack as a parameter, the lift coefficient can be
expanded to the following:
\beq
C_L = C_{L\alpha}(\alpha-\alpha_0)
\eeq
The parameter $\alpha_0$ is the angle of attack that results in zero
lift. This is a function of the airfoil shape. The parameter
$C_{L\alpha}$ is the lift curve slope and is a function of airfoil
shape, wing shape and Reynolds number. In order to remove, wing shape
from the design, the aspect ratio is used to convert the wing lift to
a sectional airfoil coefficient.
\beq\label{e:ar_equation}
C_{L\alpha} = \frac{C_{l\alpha}}{1+\frac{C_{l\alpha}}{\pi e AR}}
\eeq
The parameter $e$ is an efficiency parameter which is often assumed to
be 80-90\%. The coefficient, $C_{l\alpha}$ is the airfoil lift curve
slope which is different than the wing lift curve slope. The wing lift curve slope
will always be smaller than the airfoil lift curve slope. This is
because of an effect called wing tip vortices. Wing tip vortices are
an aerodynamic effect where high pressure from the bottom of the wing
moves around the wingtips to the area of low pressure. The only way to
mitigate these vortices is by installing winglets or increasing the
aspect ratio. Note that winglets increase drag and weight which is why
you don't typically see them on RC aircraft. The coefficient
$C_{l\alpha}$ is then simply a function of airfoil shape and
Reynolds number. This is what is often referred to as 2D lift. It is
the lift of airfoils which are two dimensional rather than 3D lift
which is over an entire wing. It is at this point that airfoil
selection and design can be used. The website \url{airfoiltools.com}
is a great resource for plotting lift curve slopes of various airfoil
shapes as a function of Reynolds number. I also have a great Youtube
video on how to use XFLR5 (pronounced X-Flyer
Five)\cite{XFLR5_tutorial}. A more comprehensive guide on XFLR5 can be
found in \cite{XFLR5_Guidelines} while the software itself can be
downloaded in \cite{XFLR5}. The basics of airfoil selection then break
down into the following 
process.

First, using your max takeoff weight and cruise flight speed, compute
the lift coefficient required in cruise. This assumes that lift equals
weight. 
\beq
C_L = \frac{2W}{\rho V^2 S}
\eeq
Then using your flight speed and chord length, compute the Reynolds number
of your aircraft. Use this Reynolds number and airfoiltools or XFLR5
to compute the sectional lift characteristics of various
airfoils. Airfoil selection can be as complicated as you make it but
you're looking for the highest lift to drag ratio airfoil. Another
simple way is to just select the airfoil with the highest sectional
lift coefficient. Another item to consider is manufacturing. Some
airfoils may be feasible for full sized aircraft but not for RC
flyers. My recommendation is to build an aircraft with a Clark-Y
airfoil. They are easy to cut with balsa or shape with foam and have
good lift to drag characteristics. Note that this airfoil is
cambered. If you want to fly upside down I suggest you use a symmetric
airfoil like a NACA 0012 or NACA 0014 is you need a bit more thickness
to fit a wing spar through the airfoil.
Once you have your airfoil selected, use the lift curve slope to
estimate $C_{l\alpha}$ by fitting a linear trend line to the portion
of the graph before stall. Also make sure to take note of what the
zero lift angle of attack is. You can then compute the wing lift curve
slope by using equation \ref{e:ar_equation}. Once you have
$C_{L\alpha}$, you can compute the angle of attack needed for
cruise. Make sure to convert $\alpha_0$ to radians before you use the
equation below.
\beq
\alpha = \frac{C_L}{C_{L\alpha}}+\alpha_0
\eeq
The resulting answer will be in radians and needs to be converted to
degrees to make sure you are not close to stall. Typically in cruise,
the angle of attack is only a few degrees with perhaps 8 degrees on
the high end. If the answer you receive is higher than that it can
mean a few things which may involve a redesign.
\begin{enumerate}[itemsep=-5pt]
  \item The simplest way to get more lift is to fly faster. The
    problem is you need a bigger motor which will increase your weight
    which will require more lift and more angle of attack. Flying
    faster will also change your Reynolds number. 
  \item You can increase the aspect ratio of your wing which will make
    your aircraft more efficient. The problem is that will increase
    the bending moment at the root creating the need for stronger
    materials at the root which also increases weight. This will also
    change your mean aerodynamic chord which will change your Reynolds
    number. 
  \item You can increase the area of the wing. This will also increase
    drag and weight but if the material you are using has a high lift
    to weight ratio then adding more wing area might be a good
    option. Depending on how you change the aircraft wing shape, the
    aspect ratio and/or the mean aerodynamic chord might change
    meaning you'll have to recompute the wing lift curve slope as well
    as the Reynolds number. 
  \item If you are using a symmetric airfoil it's possible you could
    forgo flying upside down and switch to a cambered airfoil which
    has more lift.
\end{enumerate}

Regardless of what you do make sure you angle of attack in cruise is
low which will reduce drag. It's optimal to fly at the aircraft's
highest lift to drag ratio but radio controlled aircraft don't
typically do that. Once you have redesigned your vehicle to fit within
tolerable limits it's time to look at some aircraft performance
characteristics which include the W/S (wing loading) and the T/W
(thrust to weight ratio). 

\subsection{Wing Loading and Thrust to Weight Ratio}

\subsection{Stability and Control, Center of Mass and Static Margin}

\subsection{Re-Design - Iteration and Detailed Sketch}

\subsection{Computer Aided Design}

\subsection{Purchase Components}

\subsection{Flying}

\bibliographystyle{unsrt}
\bibliography{../papers.bib}

\newpage

\input{nomenclature}

\end{document}
