%\document class[]{aiaa-tc} %load in aiaa class template file
\documentclass{article}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{mathabx}
\usepackage{hyperref}
\usepackage{natbib} %REMOVE THIS IF YOU DON'T WANT BRACKETS ON YOUR PAPER
\usepackage{float}
\usepackage{color,soul}
\usepackage[font=normalsize,labelfont=bf]{caption}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}

\newcommand{\sref}[1]{$^{[\ref{#1}]}$}
\newcommand{\dref}[2]{$^{[\ref{#1}-\ref{#2}]}$}

 % Define commands to assure consistent treatment throughout document
\newcommand{\eqnref}[1]{(\ref{#1})}
\newcommand{\class}[1]{\texttt{#1}}
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\file}[1]{\texttt{#1}}
\newcommand{\BibTeX}{\textsc{Bib}\TeX}
%\newcommand{\href}[1]{\href{#1}{

\hypersetup{%
  colorlinks=true,% hyperlinks will be black
  linkcolor=blue,
  linkbordercolor=red,% hyperlink borders will be red
   %pdfborderstyle={/S/U/W 1}% border style will be underline of width 1pt
}

\makeatletter
\Hy@AtBeginDocument{%
  \def\@pdfborder{0 0 1}% Overrides border definition set with colorlinks=true
  \def\@pdfborderstyle{/S/U/W 1}% Overrides border style set with colorlinks=true
                                % Hyperlink border style will be underline of width 1pt
}
\makeatother

\begin{document}

\title{Project Based Engineering Instrumentation With CircuitPython}
\author{Carlos Montalvo\footnote{Montalvo: William B. Burnsed Jr. Department of Mechanical, Aerospace and Biomedical Engineering, Associate Professor}
  \\University of South Alabama\\Mobile, AL\ \\ \ \\Marine Feron Leabeater \footnote{Leabeater: Lounsbury College of Education, EdD Graduate Student}\\Georgia College and State University \\ Milledgeville, GA\ \\ \ \\Lisa Schibelius\footnote{Schibelius: Department of Engineering Education, PhD Graduate Research Assistant}\\
Virginia Polytechnic Institute and State University\\Blacksburg, VA}

\maketitle

\linespread{1}

\newpage

\section*{Current Edition}

This manuscript was last updated on \today. 
The latest edition can be found on \href{https://github.com/cmontalvo251/LaTeX/blob/master/PBL_CircuitPython_Instrumentation/main.pdf}{Github}

\section*{Manuscript Changes}

\begin{enumerate}[itemsep=-5pt]
\item Original tutorials in Google Docs created
\item December 21st, 2021 - Updated links for manuscript and hardware
\item Tutorials purchased by \href{https://tangibles-that-teach.gitbook.io/instrumentation-lab-manual/-MbMx70LQzRmEG_hS7Ld/}{Tangibles that Teach}
\item May 30th, 2022 - Tangibles that teach went out of business and
  chapter began the move to Github
\item June 28th, 2022 - Work began on a Chromebook. Unfortunately the
  Figures folder is not back up on Git. As such a main\_latest.pdf has
  been created that's the latest full version. The main.pdf is the
  version created by the Chromebook so it has new chapters but none of
  the older chapters. Figure files are now backup on Git but only
  figures from the 'Voltage Potentiometer' are currently there.
\item July 2nd, 2022 - All figures backed up and latest manuscript completed
\item October 18th, 2022 - A pedometer lab has been added.
\item November 8th, 2022 - Edited the Servo and feedback control servo
  lab to be one big lab with 3 parts
\item January 8th, 2023 - Removed all mentions of Tangibles that Teach
  in the main body of the text.
\item February 6th, 2023 - Changed an assignment description for LEDs
  and push buttons
\item February 22nd, 2023 - Moved the Bluetooth module to be before
  the modules lab so this will end up being a few updates to make
  those projects more uniform. I also created a backup called
  Bluetooth original in case you wanted to go back to the other version.
\item June 16th, 2023 - Cleaned up the ``changes needed" list and updated Method 3 with some new software updates.
\item August 9th, 2023 - Edited the preamble for assignments and made a note in the servo lab
\item September 5th, 2023 - Updated hyperlinks to show up on Microsoft Edge using underlines and different color text.
\item September 7th, 2023 - Fixed the TL;DR section
\item September 15th, 2023 - Updated a link in the DAQ lab
\item September 20th, 2023 - Made the modules installation section a standalone tex file
\item October 4th, 2023 - Added a new potentiometer photo
\item November 16th, 2023 - Edited the lists of parts 
\item November 22nd, 2023 - Edited acceleration lab
\item December 5th, 2023 - Edited the troubleshooting section
\item January 12th, 2024 - Edited chapters 1 and 2
\item January 22nd, 2024 - Edited the bootloader update notes in the troubleshooting chatper
\item January 25th, 2024 - More edits to the bootloader notes
\item March 20th, 2024 - Added a preamble about where this textbook is located
\item April 27th, 2024 - Previously edited servo lab and pendulum lab just the assignment portion and then added this change log to the manuscript
\item May 7th, 2024 - Title page changed and many sections moved around. Added a new results and discussion section following data analysis from course surveys of this course.
\item August 12th, 2024 - Added a new requirement for the servo lab
\item September 12th, 2024 - A few edits to assignment descriptions
\item October 25th, 2024 - Updated the photocell lab to include the number of data points for the histogram
\item November 11th, 2024 - Updated Method 3 quick list
\item March 26th, 2025 - Updated servo, pitot probe and photocell assignments
\item May 24th, 2025 - Made a few changes to chapters 2 and 19 to transition between the projects and the results and discussion about lab at home frameworks and project based teaching. Also overhauled the feedback control project. 
\item June 16th, 2025 - Updated the assignment directions for all projects.
\item June 17th, 2025 - Added more theory to the photocell, thermistor and pitot probe assignment
\item June 20th, 2025 - Added more figures and wiring diagrams to many assignments to help students wire up their circuits.
\item June 21st, 2025 - Added a few problems from a standard instrumentation textbook. Probably need more statistics questions
\item July 2nd, 2025 - Moved changes needed to the preamble. Added learning objectives and parts list for a few projects that had them missing.
\item September 26th, 2025 - Updated the potentiometer assignment to include converting voltage to angle in degrees.
\item October 1st, 2025 - Updated the thermistor modules assignment with the thermistor equations so it's easy to find in two places.
\item November 12th, 2025 - Fixed an equation, link and figure in the feedback control assignment
\end{enumerate}

\section*{Changes Needed}
\begin{enumerate}[itemsep=-5pt]
\item Move all of these changes over to Github issues for better tracking
\item More theory is needed or direction to further reading 
\item The circuit photo for the LSM6DS33 is not correct (SDA and SCL are backwards)
\item Figure captions to all figures
\item Proper citations to other websites
\item More statistics questions
\item Add labels describing components of the kit in the main kit picture
\end{enumerate}

\section*{Future Project Ideas}
\begin{enumerate}[itemsep=-5pt]
\item PhotoInterrupter Lab (Angular Velocity) - 3D print a disc with holes on the outside to eventually mount to a rotating shaft  
\item Load Cell Calibration Lab - load cells with the HMC converter
\item Sunlight/Clouds - Take temperature and light data on a cloudy day. Draw correlations between sunlight and temperature
\item Sunrise/Sunset - Take temperature and light data over the course of a whole day
\end{enumerate}

\newpage

\tableofcontents

\newpage

\input{intro}

\input{projects}

\input{results_discussion}

\input{acknowledgements}

\bibliographystyle{unsrt}
\bibliography{../papers.bib}

\end{document}
