\section{Troubleshooting Guide}

{\bf Is your CPX/CPB broken or not running code? Read below.}

\begin{enumerate}[itemsep=-5pt]
  \item Your CPX is a USB stick that can hold as many Python files as 2MB will allow but it can only run or execute one python script at a time. It will look for code.py and then main.py and that's it. 
  \item Make sure you're using Mu in the right Mode (CircuitPython)
  \item Make sure main.py or code.py is on the CIRCUITPY drive and not somewhere on your computer.
  \item Make sure you are editing the right file in Mu. Do you have two versions of main.py or perhaps main.py and code.py?
  \item Do you have boot.py on there when you don't need it?
  \item Are you editing using Thonny or Spyder? You're supposed to use Mu. 
  \item Are you editing a file on your computer? Make sure you are
    writing to the CIRCUITPY drive.
  \item Do you have the right modules in your {\it lib} folder?
  \item Unplug the CPX, close Mu and try again.
  \item First, try and reset the CPX to CPLAYBOOT and reflash the UF2 to see if that fixes it.
  \item If you have a linux computer you can ``sudo screen /dev/ttyACM0" and then run ``import storage" and then ``storage.erase\_filesystem()"
  \item If that doesn’t work sometimes you just need to completely erase the CIRCUITPY drive so head over to this \href{https://learn.adafruit.com/adafruit-circuit-playground-express/troubleshooting}{troubleshooting guide} and follow some of the steps they tell you.
  \item As a last resort you can try to download these UF2s and
    hopefully it will fix all errors and mistakes -
    \href{https://cdn-learn.adafruit.com/assets/assets/000/048/745/original/flash_erase_express.ino.circuitplay.uf2?1512152080}{CPX}, \href{https://cdn-learn.adafruit.com/assets/assets/000/082/950/original/CP_Bluefruit_QSPI_Erase.UF2?1572026649}{CPB}
  \item If you're having an issue with an old bootloader be sure to update the \href{https://learn.adafruit.com/adafruit-circuit-playground-bluefruit/update-bootloader-use-command-line}{bootloader for the CPB} or alternatively the \href{https://learn.adafruit.com/adafruit-circuit-playground-express/updating-the-bootloader}{bootloader for the CPX}. At the time of this writing here is the command you need to upload the new bootloader the CPB.
  \begin{verbatim}adafruit-nrfutil --verbose dfu serial --package 
circuitplayground_nrf52840_bootloader-0.8.2_s140_6.1.1.zip -p /dev/ttyACM0 
-b 115200 --singlebank --touch 1200\end{verbatim}
  For the CPX there is a UF2 that you transfer over to the CPLAYBOOT drive. At the time of this writing the version you're looking for is {\it update-bootloader-circuitplay\_m0-v3.15.0.uf2}.
  \item Are you running out of memory? To figure out how much RAM you have on your CPX/CPB you need to {\it import gc} then run {\it gc.collect()} and then {\it print(gc.mem\_free())}. At the time of this writing, the CPX has about 17 KB of RAM and the CPB has around 140 KB of RAM. 

\end{enumerate}
