%\document class[]{aiaa-tc} %load in aiaa class template file
\documentclass{article}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{mathabx}
\usepackage{hyperref}
\usepackage{natbib} %REMOVE THIS IF YOU DON'T WANT BRACKETS ON YOUR PAPER
\usepackage{float}
\usepackage{multicol}
\usepackage{color,soul}
\usepackage[font=normalsize,labelfont=bf]{caption}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}

\hypersetup{%
  colorlinks=true,% hyperlinks will be black
  linkcolor=blue,
  linkbordercolor=red,% hyperlink borders will be red
   %pdfborderstyle={/S/U/W 1}% border style will be underline of width 1pt
}

\newcommand{\sref}[1]{$^{[\ref{#1}]}$}
\newcommand{\dref}[2]{$^{[\ref{#1}-\ref{#2}]}$}

 % Define commands to assure consistent treatment throughout document
\newcommand{\eqnref}[1]{(\ref{#1})}
\newcommand{\class}[1]{\texttt{#1}}
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\file}[1]{\texttt{#1}}
\newcommand{\BibTeX}{\textsc{Bib}\TeX}
\def\beq{\begin{equation}}
\def\eeq{\end{equation}}
\def\beqn{\begin{matrix}}
\def\eeqn{\end{matrix}}

\begin{document}

\begin{center}
\begin{LARGE}{\bf Aerospace Mechanics and Controls}\end{LARGE}\\
\large
\vspace{22 mm}
   A Brief Textbook Presented to the \\ 
   Student Body of the University of South Alabama \\
\vspace{22 mm}
%%by\\
\vspace{22 mm}
  %%Carlos Montalvo \\
\vspace{22 mm}
       %%{\itshape University of South Alabama}\\
Last Update: \today\\
{\bf Copyright $\copyright$ Carlos Jos\'{e} Montalvo}
\end{center}

\linespread{1}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{}

\section*{Current Edition}

This manuscript was last updated on \today. 
The latest edition can be found on \href{https://github.com/cmontalvo251/LaTeX/blob/master/Aerospace_Mechanics/aerospace_mechanics.pdf}{Github}

\section*{Manuscript Changes}

\begin{enumerate}[itemsep=-5pt]
\item June 10th, 2020 - Magnetic Field Model section updated to
  reflect the difference between the East North Vertical and the North
  East Down reference frames. The Figure showing the magnetic field
  for an example Low Earth Orbit has also been update. Also added this
  page for manuscript changes and the following pages that list where
  this file can be found.
\item December 10th, 2020 - Moved to public Github Repo separate from
  MATLAB
\item December 30th, 2020 - Moved papers.bib to parent directory Added
  datetime to title page. Added  section headers to RC aircraft
  design. Wrote text all the way through the airfoil selection. Added
  a references section.
\item December 31st, 2020 - Finished RC aircraft section.
\item September 30th, 2021 - Renamed report to Aerospace Mechanics and
  Controls
\item December 21st, 2021 - Moved CubeSAT abstract to introduction on
  CubeSATs. Imported entire Aircraft Mechanics textbook into here
\item December 22nd, 2021 - Added a section on GNC design for
  CubeSats. Added acknowledgements section.
\item June 2nd, 2022 - Added some items to changes needed and fixed
  two references
\item July 30th, 2022 - Included a derivation of direct measurement of Euler Angles using an IMU.
\item July 31st, 2022 - Added GPS coordinate conversion to cartesian coodinates as well as heading angle and speed estimation from GPS.
\item October 27th, 2022 - Added computation of lat, lon, alttiude to ECI frame.
\item March 20th, 2024 - Added a Current Edition section above
  Manuscript Changes. Also added color to hyperlinks
\item May 7th, 2024 - Added a battery sizing section to the aircraft
  design chapter
\item July 20th, 2024 - separated sections into separate files...Also started adding the quacopter aerodynamic model
\item July 22nd, 2024 - Finished the quadcopter aerodynamics section and made a quick edit to the GNC aircraft PID control scheme
\end{enumerate}

\section*{Changes Needed}

\begin{enumerate}[itemsep=-5pt]
  \item Aircraft Changes
    \begin{enumerate}[itemsep=-5pt]
    \item Include some plots on RC aircraft design
    \item Add an example RC aircraft design
    \end{enumerate}
\item Spacecraft Changes
  \begin{enumerate}[itemsep=-5pt]
  \item Add some work on Kerbal Space Program
  \item Need to make Orbital Elements its own section. Discuss how to
    get position and velocity from orbital elements and orbital elements
    from position and velocity.
  \item Need to explain the difference between Geodetic and Geocentric
    coordinates
  \item Need to add the computation of latitude an longitude from
    cartesian and vice versa as well as flat Earth approximation
  \item Derivation of ground path taking into account orbital
    precession, rotation of the Earth and swath angle from projection of
    a satellite onto the Earth. See pdf from the Air Force. A Google
    search will hopefully turn up the paper I'm thinking of.
  \item Consider adding the section on pointing analysis
  \end{enumerate}
\item General Changes
  \begin{enumerate}[itemsep=-5pt]
  \item Add a section on PBL of a high powered rocket
  \item Add an abbreviations section
  \item Add Project Based learning to this manuscript
  \item Direct the reader to my Instrumentation textbook to build a
    datalogger to put on an airplane or rocket.
  \item Direct reader to FASTkit to run dynamics in simulation
  \item Include some simulation results of aircraft, rocket and
    satellite from FASTkit.
  \item Need to add derivation of a complimentary filter using
    transfer functions which means I probably need to add in transfer
    functions into this manuscript
  \item Need to finish discrete dynamics section
  \item Need example ID and ADs
  \item Need to add parallel axis theorem for inertia computation
  \item Add GNC for all vehicles: spacecraft, rocket, aircraft, quadcopter
  \end{enumerate}
\end{enumerate}

\section*{Acknowledgements}

Carlos Montalvo would like to thank numerous students for their
contribution to this document. They have been instrumental in making
this textbook a reality and this textbook would not be where it is
today without them. Those students are: Weston Barron, Colin Mcgee,
Darcey D'Amato, Ruthie Hill, Drew Russ, William Sherman, Maxwell Cobar, Wei Min Patrick, Caroline
Franklin, Andrew Givens, Aaron Godfrey, Nghia Huynh, Lisa Schibelius,
and Brandon Troub.

\newpage

\tableofcontents

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{introduction}
\input{nomenclature}
\input{particle_dynamics}
\input{rigid_bodies}
\input{eoms}
\input{external_models}
\input{forces_moments}
\input{gnc}
\input{sys_eng}
\input{cubesats}
\input{aircraft_design}
\input{numerical_integration}
\input{aircraft_design_equations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{unsrt}
\bibliography{../papers.bib}

\end{document}
