\section{Introduction}

This report represents aerospace mechanics and controls for CubeSats,
quadcopters and aircraft. A CubeSAT is a small satellite on the order
of 10 centimeters along each axis. A 1U satellite is a small cube with
10 cm sides. These satellites are used for a variety of missions and
created by a variety of different organizations. When deployed from 
a rocket, a CubeSAT may obtain a large angular velocity
which must be reduced before most science missions or communications
can take place. Maximizing solar energy charging also involves
better pointing accuracy. To control the attitude of these small
satellites, reaction wheels, magnetorquers and even the gravity
gradient are used in low earth orbit (LEO) while reaction control thrusters
are typically used in deep space.  On a standard LEO CubeSAT, 3 reaction
wheels are used as well as 3 magnetorquers. In the initial phase of
the CubeSAT mission, the magnetorquers are used to reduce the angular
velocity of the satellite down to a manageable level. Once the norm of the angular velocity is
low enough, the reaction wheels can spin up reducing the angular
velocity to zero. At this point a Sun finding algorithm is employed to
find the Sun and fully charge the batteries. In LEO two independent
vectors are obtained, the Sun vector and the magnetic field vector, to
determine the current attitude of the vehicle which is typically
called attitude determination. Other sensors such as horizon sensors,
star trackers and even lunar sensors can be used to obtain the
quaternion of the vehicle. This paper investigates the necessary
mathematics to understand the intricacies of guidance, navigation and
control specifically discussing the attitude determination and
controls subsystem (ADACS).